---
title: "ML Project"
subtitle: "Complementary Code"
author: "Daniel Choy, Yi Cao, Danny Zeng"
date: "Nov 10, 2020"
output:
  html_notebook:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Machine Learning Project - Ames Housing Data

Ames, Iowa is the college town of **Iowa State University**. The Ames housing dataset consists of about $2500$ house sale records between $2006-2010$. Detailed information about the house attributes, along with the sale prices, is recorded in the dataset. The goal of the project is to:
- perform descriptive data analysis to gain business (i.e. housing market) insights
- build descriptive machine learning models to understand the local housing market.
- build predictive machine learning models for the local house price prediction.


### Who We Are

Choose (but are not limited to) one of the following options:
- You are a housing market consultant.
- You are a data scientist for an online real estate database company which provides house price estimation (like **Zillow** home price estimates) for the site visitors.

The goal is to build a highly accurate predictive or descriptive model. To accomplish this, it is important to understand the local housing market in a **data driven** way. 

### Import Libraries

```{r, echo=False, }
library(dplyr)
library(ggplot2)
library(lubridate)
library(tidyverse)
```

### Load Data

There are two data sets included in the data folder: Ames_Housing_Price_Data.csv and Ames_Real_Estate_Data.csv.

The Ames_Housing_Price_Data.csv set contains  81  data columns, including the key feature SalePrice which will be used as the target of the predictive/descriptive modeling. PID refers to the land parcel ID, which can merged on the MapRefNo column of the Ames Accessor Data (Ames_Real_Estate_Data.csv) to find the property address. Using a free service, such as geopy, we can find the long-lat coordinates of the houses.

The columns of the data are mostly attributes associated with the land and the houses. There are size related attributes, quality and condition attributes, house attachment attributes, etc.

```{r}
housing = read.csv('./Ames_HousePrice.csv', header = TRUE)
```

### Data Inspection
```{r}
colnames(housing)
```

```{r}
str(housing)
```

### Data PreProcessing

```{r}
housing_df = housing %>% 
  
  # filter out NA's and duplicates
  filter(is.na(MasVnrArea) != 1) %>%
  filter(is.na(TotalBsmtSF) != 1) %>%
  filter(PID != 909276070 & X != 1) %>%
  
  # Create new columns for better modeling   
  mutate(
    
    MasVnrArea2 = case_when(
      MasVnrArea == 0 ~ 0,
      MasVnrArea != 0 ~ 1),
    
    LotFrontage2 = ifelse(is.na(LotFrontage), 0, LotFrontage),
    
    Alley2 = ifelse(is.na(LotFrontage), 0, 1),
    
    LotShape2 = ifelse(LotShape == 'Reg', 'Regular', 'Irregular'),
    
    GarageArea2 = ifelse(is.na(GarageArea), 0, GarageArea),
    
    PoolArea2 = ifelse(is.na(PoolArea), 0, PoolArea),
    
    PoolArea3 = ifelse(is.na(PoolArea), 0, 1),
    
    X2ndFlrSF = ifelse(is.na(X2ndFlrSF), 0, X2ndFlrSF),
    
    HalfBath2 = case_when(
      HalfBath == 1 ~ 0.5,
      HalfBath == 0 ~ 0),
    
    BsmtHalfBath2 = case_when(
      BsmtHalfBath == 1 ~ 0.5,
      BsmtHalfBath == 0 ~ 0),
    # If Basement Finish Type1 is GLQ or ALQ, we use the Basement Finish Square Feet1
    BasmtFinSF1 = ifelse(BsmtFinType1 == 'GLQ' | BsmtFinType1 == 'ALQ', BsmtFinSF1, 0),
    # If Basement Finish Type2 is GLQ or ALQ, we use the Basement Finish Square Feet2
    BasmtFinSF2 = ifelse(BsmtFinType2 == 'GLQ' | BsmtFinType2 =='ALQ', BsmtFinSF2, 0),
    # The Total Square Feet of Basement Finish is BasmtFinSF1 + BasmtFinSF2
    BasmtFinSF = BasmtFinSF1 + BasmtFinSF2,
    # Total Living Area is GrLivArea + BasmtFinSF
    total_LivArea = GrLivArea + BasmtFinSF,
    num_bathroom = BsmtHalfBath2 + HalfBath2 + FullBath + BsmtFullBath
  )

# export to csv
write.csv(housing_df, file = 'housing_v2.csv')

```
